Propositions as Sessions, Mechanically

### 2015-02-24 paper/paper.tex
Tweaking content further. Hyperref and bibliography acting up.
Reference overlaps page boundary causing hyperref to explode. I
need to investigate this further.

\usepackage{macros}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{listings}
\usepackage{lstcoq}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cmll}
\usepackage{prooftree}
\usepackage{amsmath}
\newcommand{\typestate}[2]{{#1}\gg{#2}}
\usepackage[style=alphabetic,sorting=nymt,natbib=true,
  maxbibnames=100]{biblatex}
\addbibresource{paper.bib}

@@

The locally nameless first-order representation for binders and free variables
was adopted for both GV and CP, following
\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM}. This first-order
variant prevents variable-capture issues when compared with other first-order
approaches. Alternatively, I could have chosen a higher-order abstract syntax
(HOAS) approach such as the parametric version (PHOAS) described
by~\citeauthor{Chlipala:2008:PHOAS}~\cite{Chlipala:2008:PHOAS}. PHOAS relies
on the metalanguage (Coq, in this case) for handling binders and substitution
so has less overhead in managing syntax than the locally nameless
approach. However, it was less clear how to handle linearity using this
approach and has the disadvantage of being able to represent terms not valid
in the language being mechanised (so called ``exotic'' terms). Further, a
higher-order approach would prevent the use of the Metatheory library which
provides a number of useful tactics for manipulating typing contexts. Previous
efforts to mechanise \fpop~\cite{Park:2014:MMW} illustrate how to handle a
language containing both linear and non-linear contexts. However, to stay true
to CP/GV, I maintained a single context. Interestingly, it became clear that
the mechanised version of CP required explicit weakening and contraction to
facilitate the reduction rules.

For handling binder freshness, I chose to adopt the cofinite quantification
approach described by~\cite{Aydemir:2008:EFM}, which excludes a finite set of
variables from being considered as the binder. In contrast, the traditional
``exists-fresh'' approach, where the binder is only required to be fresh
within the abstraction's body, does not produce strong enough induction
principles in some cases. This change will necessitate changes to the typing
rules of GV and CP. Consider for instance the Connect rule in
GV~\cite{Wadler:2014} (C) and its corresponding cofinite quantification
version (CCO): \input{with-connect-rule}

The conclusion specifies a variable in (C) but a type in (CCO) (de Bruijn
index rather than a name). Thus, in the premise for (CCO) one can choose a
suitably fresh $x$ in each premise. Note that, $L$ can be chosen such that
$\Psi \cup \Phi \cup \fv(M) \cup \fv(N) \subseteq L$, where $\fv(M)$ are the
set of free variables in term $M$, and $x$ can be the same in both
premises. Similar changes are made to the rules for CP (see
\S~\ref{sec:cp}). These changes do not alter the typing of processes nor the
semantics of the translation from GV to CP. An equivalence between
``exists-fresh'' and cofinite quantification definitions of the simply-typed
$\lambda$-calculus is given by
\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM}, and I suspect a similar
result could be derived straightforwardly (but tediously) for the CP/GV
system.

@@

Formalising the process calculus CP provided many challenges for the
metatheory library since to my knowledge, the library has never been applied
to a process calculus before. So for instance, in the lambda calculus one
typically does not work up to permutation of binders, whereas in a process
calculus these kinds of structural equivalences are standard.

@@

@@

I describe the formalisation as it currently stands in this section. The use
of the metatheory library from UPenn provided useful tactics and properties
regarding freshness of free variables and environments.

\subsection{Future Work}

@@

\printbibliography

>>>>

\usepackage{macros}
\usepackage{comment}
\usepackage{listings}
\usepackage{lstcoq}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cmll}
\usepackage{prooftree}
\usepackage{amsmath}
\newcommand{\typestate}[2]{{#1}\gg{#2}}
\usepackage[style=alphabetic,sorting=nymt,natbib=true,
  maxbibnames=100]{biblatex}
\addbibresource{paper.bib}

@@

\subsection{Preliminaries}\label{sec:approach}

The locally nameless first-order representation for binders and free variables
was adopted for both GV and CP, following
\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM}. This first-order
variant prevents variable-capture issues when compared with other first-order
approaches. Alternatively, I could have chosen a higher-order abstract syntax
(HOAS) approach such as the parametric version (PHOAS) described
by~\citeauthor{Chlipala:2008:PHOAS}~\cite{Chlipala:2008:PHOAS}. PHOAS relies
on the metalanguage (Coq, in this case) for handling binders and substitution
so has less overhead in managing syntax than the locally nameless
approach. However, it was less clear how to handle linearity using this
approach and has the disadvantage of being able to represent terms not valid
in the language being mechanised (so called ``exotic'' terms). Further, a
higher-order approach would prevent the use of the Metatheory library which
provides a number of useful tactics for manipulating typing contexts.

Previous efforts to mechanise \fpop~\cite{Park:2014:MMW} illustrate how to
handle a language containing both linear and non-linear contexts. However, to
stay true to CP/GV, I maintained a single context. Interestingly, it became
clear that the mechanised version of CP required explicit weakening and
contraction to facilitate the reduction rules (\S~\ref{sec:cp}).

\input{with-connect-rule}

For handling binder freshness, I chose to adopt the cofinite quantification
approach described by~\citeauthor{Aydemir:2008:EFM}, which excludes a finite
set of variables from being considered as the binder. In contrast, the
traditional ``exists-fresh'' approach, where the binder is only required to be
fresh within the abstraction's body, does not produce strong enough induction
principles in some cases. This change will necessitate changes to the typing
rules of GV and CP. Consider for instance the Connect rule in GV (C) and its
corresponding cofinite quantification version (CCO) in
Figure~\ref{fig:connect}. The conclusion specifies a variable in (C) but a
type in (CCO) (de Bruijn index rather than a name). Thus, in the premise for
(CCO) one can choose a suitably fresh $x$ in each premise. Note that, $L$ can
be chosen such that $\Psi \cup \Phi \cup \fv(M) \cup \fv(N) \subseteq L$,
where $\fv(M)$ are the set of free variables in term $M$, and $x$ can be the
same in both premises. Similar changes are made to the rules for CP. These
changes do not alter the typing of processes nor the semantics of the
translation from GV to CP. An equivalence between ``exists-fresh'' and
cofinite quantification definitions of the simply-typed $\lambda$-calculus is
given by \citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM}, and I suspect
a similar result could be derived straightforwardly (but tediously) for the
CP/GV system.

@@

Formalising the process calculus CP provided many challenges for the
metatheory library since to my knowledge, the library has never been applied
to a process calculus before. So for instance, in the lambda calculus one
typically does not work up to permutation of binders, whereas in a process
calculus these kinds of structural equivalences are standard. Additionally,
substitution and opening are only defined on processes for names. In other
words, one cannot substitute an arbitrary process term for a name, unlike in
the lambda calculus. Thus frameworks which assume a lambda calculus style
substitution operation are not immediately amenable to the CP
setting~\cite{Lee:2012}.

@@

Explicit weakening and contraction were required to differentiate the
reduction rules for servers and clients. In contrast, \citeauthor{Wadler:2014}
defines these operations explicitly.

@@

\section{Future Work}

@@

%\printbibliography
